---
title: "FAO aquastat ETo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{FAO_aquastat_ETo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Download FAO climate data

The goal of this vignette is to access and download FAO climate data, then prepare that data to calculate ETo. The methods used in this package `fao56` are pulled from the FAO 56 report on evapotranspiration. Many of the functions correspond directly to an equation in that report.

```{r setup}
library(fao56)
library(dplyr)
```

Let's say we want climate data for Kuala Lumpur, which has the coordinates 3.1390, 101.6869. We first use `get_aquastat_climate()` to visit the those coordinates and seek to download climate data. On the page, click the download CSV button. It also helps to note the elevation, which is 49 m in this case.

```{r eval=FALSE}
get_aquastat_climate(lat = 3.1390, lon = 101.6869)
```

Next, move the file to a convenient locaiton. This package already has the file installed under extdata. It can be accessed by finding the location using `system.file()`. The file can then be read using `read_et0_csv()`

```{r}
kl_path <- system.file("extdata","ET0 - 3.14-101.69.csv", package = "fao56")
kl_climate <- read_et0_csv(kl_path)
kl_climate
```

## Prepare the data for FAO penman monteith

The FAO Penman-Monteith equation requires specific variables to calculate ETo. The function `gen_ETo_predictors_from_aquastat` allows calculation of these variables and the ability to select particular ones. It also calculates ETo in the field `ETo_calc`.

```{r}
z = 49
lat = 3.139
col_select <- c("Rn_MJ_per_day", "G_MJ_per_day", "Tmp_Mean_degC", "Wind_2m_m_per_s", "es_kPa", "ea_kPa", "ETo_calc", "ETo_mm_per_d")
ETo_df <- gen_ETo_predictors_from_aquastat(kl_climate, lat = lat, z = z, col_select = col_select) 
ETo_df
```

There may be small discrepancies between `ETo_calc` and `ETo_mm_per_d`. These should be small, on the order of 1%. 

```{r}
ETo_df %>% mutate(pct_error = round(100 * (ETo_calc - ETo_mm_per_d) / ETo_mm_per_d, 1)) %>%
  dplyr::select(ETo_calc, ETo_mm_per_d, pct_error)
```


